---
title: "Loxodonta africana HSM (Part IV)"
author: "Jessica Bernal"
date: '2022-08-08'
output: html_document
---

### 4. Multicollinearity analysis and variable selection 


```{r}
#setwd if neccesary. Here we only check it, as it has been previously defined for the project
getwd()
```



Loading the libraries to use,

```{r}
library(raster)
library(rgdal)
library(spatstat)
library(spatialEco)
library(usdm) # Analisis de incertidumbre para modelos de distribucion de especies
library(ggplot2) # Powerful graphics language for creating elegant and complex plots
library(maptools)
```


Loading the data and creating the paths to use. 




```{r}
### Rasters path
path2layers<-"./Loxodonta_africana/2_WorldClim_R/wc2.1_tiles"

### Results path 
path2work1<-"./Loxodonta_africana/3_test_multicolinealidad/resultados/"

### Path for selected variables
path2work2<-"./Loxodonta_africana/3_test_multicolinealidad/resultados/variables_seleccionadas/"

### Loading variables
VAR_PN <- stack(list.files(path=path2layers,pattern='',full.names=TRUE))
names(VAR_PN)
```


Crop raster with AOI

```{r}
### AOI path
PN_shp <- readOGR("./PN_Limpopo/limites_limpopo/PN_Limpopo_30km.shp")

### crs transform
PN_shp <- spTransform(PN_shp, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))

## Crop and mask with AOI
#VAR_PN[] <- 1:length(VAR_PN)
VAR_PN <- crop (VAR_PN, extent(PN_shp))
VAR_PN <- mask(VAR_PN, PN_shp)

## Check that it worked
plot(VAR_PN)
plot(PN_shp, add=TRUE, lwd=2)
```

Let`s rename the raster layers

```{r}
names(VAR_PN) <- c('bio1', 'bio2', 'bio3', 'bio4', 'bio5', 'bio6', 'bio7', 'bio8', 'bio9', 'bio10', 'bio11', 'bio12', 'bio13', 'bio14', 'bio15', 'bio16', 'bio17', 'bio18','bio19', 'alt' )

names(VAR_PN)
```

```{r}
plot(VAR_PN)
```


Descriptive statistics of the variables
```{r}
summary(VAR_PN)
```
```{r}
VAR_PN <- stack(VAR_PN)
names(VAR_PN)
```
```{r}
summary(VAR_PN)
```


Exploring the structure of variables
```{r}
print(VAR_PN)
```


```{r}
# Assign coordinate system, if necessary
#crs(VAR_PN) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
```


For the treatment of environmental variables, and in order to reduce the level of uncertainty in our predictions, the variables bio 3, bio 14 and bio15 will be excluded from the models, following the recommendations of Varela et al. (2015), given that these variables have shown that entail a high level of discrepancy between the general circulation models.

```{r}
### Remove bio_3, bio_14, bio_15 (Varela_et_al_2015_PLoSONE)
names(VAR_PN)
VAR_PN_DEF <- dropLayer(VAR_PN, c(3, 14, 15))
plot(VAR_PN_DEF)
```

We now calculate VIF based on the functions vifcor or vifstep

```{r}
v1 <- vifcor(VAR_PN_DEF, th=0.8) # Search for a pair of variables that have the maximum linear correlation (greater than th) and exclude one of them that has greater VIF. The procedure is repeated until there is no variable with a high correlation coefficient (greater than the threshold) with other variables.
v1
```



```{r}
v2 <- vifstep(VAR_PN_DEF, th=3) # Calculate the VIF for all variables, exclude one with the highest VIF (greater than the threshold), repeat the procedure until there are no variables left with VIF greater than th.
v2
```


```{r}
re1 <- exclude(VAR_PN_DEF,v1)
re1

plot(re1)
```



```{r}
re2 <- exclude(VAR_PN_DEF,v2)
re2

plot(re2)
```


Even if the variables with less collinearity are selected, they could really be correlated, since the VIF does not test well the correlations between pairs of variables

Let's calculate VIF "by hand"


```{r}
### Transform the rasters into a table
var.df<-as.data.frame(VAR_PN_DEF)

### Remove null values
var.df<-na.omit(var.df)

### Calculate the Spearman correlation coefficient between variables
var.cor<-cor(var.df, method = c("spearman"))
```


Exploring the correlation results
```{r}
print(var.cor)
```

Correlation plot

```{r}
library(corrplot) # Graphical display of a correlation matrix, confidence interval
corrplot_number<-corrplot(var.cor, method = "number")
```


```{r}
corrplot_circle<-corrplot(var.cor, type = "upper", method = "circle") # More intuitive
```


We now transform correlations into a dataframe and Explore the structure of it,

```{r}

cor.df<-as.data.frame(var.cor)

head(cor.df)
```

We export the dataframe (only the upper diagonal)

```{r}
lower<-var.cor
lower[lower.tri(var.cor, diag=TRUE)]<-""
lower.df<-as.data.frame(lower)
lower.df

write.csv(lower.df, paste(path2work1, "var_correlations.csv"),sep=",")
```

Transform the correlation matrix into a distance matrix. This test is equivalent to a correlation test but the ramp of values change. 

```{r}
var.dist <- abs(as.dist(cor.df))
```



Calculate the dendrogram based on distances (less distance = more correlation)

```{r}
var.cluster <- hclust(1-var.dist)
```



Plot of the dendrogram. What we see is the correlation distance, inverse to the correlation coefficient. 

```{r}
plot(var.cluster)
```



Select variables with a correlation < 0.8

```{r}
plot(var.cluster)
abline(h=0.2, lty=2, lwd=2, col="red")
```


We explore the plot and select uncorrelated variables.

```{r}
var.selected <- c("bio10","bio12", "bio13", "alt")
```


We build a new table with the selected variables

```{r}
var.df2<-var.df[ , var.selected]

var.cor<-cor(var.df2, method = c("spearman"))
corrplot_number<-corrplot(var.cor, type = "upper", method = "number")

```

```{r}
corrplot_circle<-corrplot(var.cor, type = "upper", method = "circle")
```


Calculate the Variance Inflation Factor (VIF)

```{r}
result.vif<-vif(var.df2)
result.vif
```


We eliminate those with the highest VIF value, as they are the most susceptible to be generating noise.

```{r}
# Example removing bio7
#var.df2$bio7<-NULL
#result.vif<-vif(var.df2)
#result.vif
```



Now we can say that the 4 remaining vars are the ones we are going to use to model the bioclimatic space.

We represent a basic horizontal boxplot,

```{r}
p <- ggplot(data=result.vif, aes(x=Variables, y=VIF), fill = activity) +
  ggtitle("VIF") +
  geom_bar(stat="identity") +
  coord_flip() +
  theme(
    plot.title = element_text(size=18, face="bold"),
    axis.title=element_text(size=18, face="bold"),
    legend.title=element_text(size=14, face="bold"),
    legend.text=element_text(size=12),
    axis.text=element_text(size=14, colour="black"))

p
```


We get the names of the variables

```{r}
var.selected<-names(var.df2)
var.selected
```


We join all the rasters in a single object

```{r}
var.def<-brick(VAR_PN[[var.selected]])
```



Variables plot

```{r}
plot(var.def)
```


We save the selected variables to calibrate our ENMs

```{r}
writeRaster(x=var.def, path2work2, names(var.def), bylayer=TRUE, format="ascii", overwrite=TRUE)
```


We remove some objects and gain RAM

```{r}
rm(var.cor, var.df, var.df2, result.vif, var.cluster, var.dist, var.selected)
gc()
```

