---
title: "Loxodonta africana HSM (Parte VII)"
author: "Jessica Bernal"
date: '2022-08-11'
output: html_document
---

### 5. Species Distribution model with biomod2


```{r}
#setwd if neccesary. Here we only check it, as it has been previously defined for the project
getwd()
```

Loading the libraries to use,

```{r}
library(raster)
library(dismo)
#install.packages("rgeos")
library(rgeos) # Geometric operations with geographic information
#install.packages("biomod2")
library(biomod2) # Species distribution and assembly models
library(ggplot2)
#install.packages("RColorBrewer")
library(RColorBrewer) # Create nice color palettes especially for thematic maps
library(dplyr)
```


Loading the data and creating the paths to use. 

```{r}
# Load species data
DataSpecies <- read.csv2('./Loxodonta_africana/1_presencias/2_filtering_records/thinned_data_thin1.csv',sep=',', dec = ".")


### Explore the data
head(DataSpecies, n = 5)

```
```{r}
### Plot
qplot(x=X,y=Y,data=DataSpecies)
```
Convert the data to a spatial object

```{r}

library(sp)
spPointData <- SpatialPoints(DataSpecies[,c("X","Y")], 
                             proj4string = CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))

```



Name of the study species
```{r}
myRespName <- 'La'
```



Presence/absence data for Loxodonta africana

```{r}
myResp <- spPointData
myResp
```


Environmental variables (predictors)

```{r}
myExpl<-stack(list.files(path="./Loxodonta_africana/4_variables_definitivas/var_def",pattern='asc',full.names=TRUE))
crs(myExpl) <- CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
plot(myExpl)
myExpl
```


Set the output directory of the results

```{r}
path2results<-"./Loxodonta_africana/5_biomod2"
```



Number of training points

```{r}
Npresences <- length(myResp)
```


Data format

We configure biomod for calibration

```{r}
# Number of PA sets (pseudoabsences) = 10
# Number of PA per set = Number of presences
# PA selection strategy = 'random' 

myBiomodData <- BIOMOD_FormatingData(resp.var = myResp,
                                     expl.var = myExpl,
                                     resp.name = myRespName,
                                     eval.resp.var = NULL,
                                     eval.expl.var = NULL,
                                     eval.resp.xy = NULL,
                                     PA.nb.rep = 10, 
                                     PA.nb.absences = Npresences, 
                                     PA.strategy = 'random',
                                     PA.dist.min = 0,
                                     PA.dist.max = NULL,
                                     PA.sre.quant = 0.025,
                                     PA.table = NULL,
                                     na.rm = TRUE)
```
Print data format
```{r}
myBiomodData
```

Plot data format
```{r}
plot(myBiomodData)

```


Calibration options

```{r}
# Consulting help
?BIOMOD_ModelingOptions
```


We choose a GAM (Generalized Additive Model)

```{r}
#We change k=4 to avoid too complex models
myBiomodOption <- BIOMOD_ModelingOptions(GAM = list(k = 4),
                                         MAXENT.Phillips = list(path_to_maxent.jar="./maxent/maxent.jar"))
print(myBiomodOption)
```

### Calibration of models

Number of evaluation rounds = 20
Data train/test percentage = (depends on the number of occurrences of the study species)
Prevalence = 0.5
Importance of variables - Number of permutations = 3
Metrics for evaluation = ROC, TSS
Full models = TRUE


```{r}
# Consulting help
?BIOMOD_Modeling
```


```{r}
myBiomodModelOut <- BIOMOD_Modeling( 
  myBiomodData, # The data includes the records of the species, the name of the species and the variables
  models = c('GLM', 'GBM','GAM','CTA','ANN',
             'SRE','FDA','MARS','RF','MAXENT.Phillips.2'), # Methods (algorithms)
  models.options = myBiomodOption,
  NbRunEval=20, 
  DataSplit=80, # % of data used for training
  Yweights=NULL,#
  Prevalence=0.5, # Weighted sum of presences is equal to the weighted sum of absences
  VarImport=1,
  models.eval.meth = c('TSS','ROC'),
  BinRoc=TRUE, # The results will also be output as binary values using the Roc curve threshold for all models
  BinTSS=TRUE, # Results will also be output as binary values using the TSS threshold for all models
  FiltRoc=TRUE, # Results will also be output as filtered values using the Roc curve threshold for all models
  FiltTSS=TRUE, # Results will also be output as filtered values using the TSS threshold for all models
  SaveObj = TRUE, # Keep all the results and outputs on the hard drive BEWARE THAT THE COMPUTER PETA
  do.full.models = TRUE, # Models calibrated and evaluated with the entire dataset
  modeling.id = "allmodels") # ID (= name) of the modeling procedure
```



Modeling summary

```{r}
myBiomodModelOut 
```



### Models evaluation

```{r}
options(max.print=999999) # Give more RAM
```


Obtain the evaluation values of the models
```{r}
myBiomodModelEval <- get_evaluations(myBiomodModelOut)
#myBiomodModelEval
```


```{r}
xx <-get_evaluations(myBiomodModelOut, as.data.frame=T)
head(xx, n = 5)

```


Print the dimnames of this object
```{r}
dimnames(myBiomodModelEval)
```



Print ROC scores of all selected models. Models which ROC scores are above 0.7 are considered good, while >0.9 very good.

```{r}
print(myBiomodModelEval["ROC","Testing.data",,,])
```



Print TSS scores for all selected models
```{r}
print(myBiomodModelEval["TSS","Testing.data",,,])
```



Obtaining boxplot stats of ROC
```{r}
print(fivenum(as.numeric(myBiomodModelEval["ROC","Testing.data",,,])))
```



Obtaining boxplot stats de TSS
```{r}
print(fivenum(as.numeric(myBiomodModelEval["TSS","Testing.data",,,])))
```


Printing variables importance

```{r}
get_variables_importance(myBiomodModelOut)

var.imp.mod.out <- get_variables_importance(myBiomodModelOut, as.data.frame=T)
head(var.imp.mod.out, n = 5)
```

 
Save variables importance

```{r}
capture.output(get_variables_importance(myBiomodModelOut),
               file=file.path(path2results, 
                              paste(myRespName,"_variables_importance.csv", sep=",")))
```


Save evaluation metrics
```{r}

evalDF.ROC <- as.data.frame(myBiomodModelEval["ROC","Testing.data",,,])
evalDF.TSS <- as.data.frame(myBiomodModelEval["TSS","Testing.data",,,])

write.csv(evalDF.ROC, file = paste(getwd(),"/",path2results,"/",myRespName,"_evalDF_ROC.csv",sep=""))
write.csv(evalDF.TSS, file = paste(getwd(),"/",path2results,"/",myRespName,"_evalDF_TSS.csv",sep=""))

```


### Ensemble modeling 

```{r}
# Assemble all partial models
# We use the "top" models at 5%
# We use ROC as the threshold metric
# Ensemble methods are: mean, median, and weighted mean

quantileThresh <- quantile(evalDF.ROC, probs=0.95, na.rm=TRUE) ##TOP 5

myBiomodEM <- BIOMOD_EnsembleModeling( 
  modeling.output = myBiomodModelOut, # Model results
  chosen.models = 'all', # Models to include when assembling
  em.by='all', # Flag defining how the models will be combined to build the ensemble models
  eval.metric = c('ROC'), # Evaluation metric used to build the ensemble models
  eval.metric.quality.threshold = quantileThresh, # It could be 0.7 IE PUT: AUC>0.7 BE CAREFUL BECAUSE IT IS NOT BETTER BY TAKING THE TOP TOPS, DEPENDING ON THE DATA WE CAN LOSE INFORMATION NECESSARY FOR THE CORRECT INTERPRETATION
  prob.mean = TRUE, # Estimate mean probabilities through predictions
  prob.cv = FALSE, # Coefficient of variation
  prob.ci = FALSE, # Confidence intervals of the prob. half
  prob.ci.alpha = 0.05, # Level of significance to estimate the confidence interval. Default = 0.05
  prob.median = TRUE, # Estimate the median
  prob.mean.weight = TRUE, # Estimate the weighted mean of the probabilities
  prob.mean.weight.decay = 'proportional' ) # Define the relative importance of the weights
```


```{r}

### Print summary                     
myBiomodEM
```



Get Evaluations for Ensemble Models

```{r}
emEvalDF <- as.data.frame(get_evaluations(myBiomodEM))
head(emEvalDF, n = 2)
```

```{r}
write.csv(emEvalDF, file = paste(getwd(),"/",path2results,"/",myRespName,"_EnsMod_evalDF_AllMetrics.csv",sep="")) ##GET THE CSV TEXT TO TABLE TO PRESENT THE MODEL, A BOXPLOT CAN BE A LOT
```



### Projection of "current" models to the geographic space

```{r}
### Projection over the study area under current conditions
myBiomodProj <- BIOMOD_Projection(
   modeling.output = myBiomodModelOut, # Model outputs
   new.env = myExpl, # Environmental Variables
   proj.name = 'current', # Name of projections
   #selected.models = 'all', # models to be projected
   selected.models = get_kept_models(myBiomodEM, model = 1), ## Models considered in the assembly and projection - Change according to criteria - 1 IS THE MEDIAN
   compress = 'gzip',
   output.format = '.grd', # Format of GIS files (also *.img)
   do.stack=TRUE)
```

Summary of created object

```{r}
myBiomodProj
```

```{r}
# Files created on the hard drive
list.files("./5_biomod2/La/proj_current")
```


```{r}
mod_projPres <- get_predictions(myBiomodProj)
presentResult <- calc(mod_projPres,fun = median) # Choose any descriptive statistics you want

plot(presentResult)
```

```{r}
writeRaster(presentResult, filename = "La_Present", format = "GTiff", overwrite = T) # THIS GIVES US OUR FIRST MAP OF ALL THE INDIVIDUAL MODELS AVERAGED TO THE MEDIAN, BUT WE CAN CALCULATE VARIABILITY THROUGH THE DEVIATION, DEPENDS ON WHAT WE WANT IT TO INFORM US
```




### Projection of ensemble models

```{r}
myBiomodEF <- BIOMOD_EnsembleForecasting( 
  EM.output = myBiomodEM,
  projection.output = myBiomodProj,
  binary.meth = c('ROC', 'TSS'),
  filtered.meth = c('ROC', 'TSS'),
  output.format = '.grd')
```


```{r}
### Print summary
myBiomodEF
```


```{r}
plot(myBiomodEF)
```


```{r}
# Files created in hard drive
list.files("./5_biomod2/La/proj_current")
```

```{r}
mod_projPresEnsemble <- get_predictions(myBiomodEF);
presentEnsembleResult <- mod_projPresEnsemble[[2]] # This is the median of the set of models
```

```{r}
plot(presentEnsembleResult)
```


```{r}
writeRaster(presentEnsembleResult, filename = "La_PresentEnsemble", format = "GTiff", overwrite = T)
```




### Models evaluation 

```{r}

### Evaluation chart by model
gg1 <- models_scores_graph( myBiomodModelOut,
                            by = 'models',
                            metrics = c('ROC','TSS') )
```

We see the performance of each model according to the metrics used (e.g. if RF calibrated models are better than SRE)



### Contribution of variables to models

```{r}
### Load models (in this case RF and GAM)
RF.mod <- BIOMOD_LoadModels(myBiomodModelOut,models='RF')
GAM.mod <- BIOMOD_LoadModels(myBiomodModelOut,models='GAM')

### Get importance of variables in this model
RF.vi <- variables_importance(get(RF.mod[1]), getModelsInputData(myBiomodModelOut,'expl.var'), nb_rand=1)
GAM.vi <- variables_importance(get(GAM.mod[1]), getModelsInputData(myBiomodModelOut,'expl.var'), nb_rand=1)

### Check the results
RF.vi$mat
GAM.vi$mat
```


```{r}
### Convert this objects to tables = data.frame
contributions_RF <- as.data.frame(RF.vi$mat)
contributions_GAM <- as.data.frame(GAM.vi$mat)
```


```{r}
### Get bar graph for RF
barplot(height = t(contributions_RF),
        beside = TRUE,
        horiz = TRUE,
        xlab = "Variable Importance RF",
        legend = c("Random Forest"))
```

```{r}
### Get bar graph for GAM
barplot(height = t(contributions_GAM),
        beside = TRUE,
        horiz = TRUE,
        xlab = "Variable Importance GAM",
        legend = c("Generalized Additive Model"))
```

### Plot of response curves

```{r}
### Load the models for which we want to extract the predicted response curves - in our case from the ensemble models
myModel <- BIOMOD_LoadModels(myBiomodEM, models='mergedData')

```


```{r}
### Plot 2D
myRespPlot2D.La <- response.plot2(models = myModel[1],
                                  Data = get_formal_data(myBiomodModelOut,'expl.var'), 
                                  show.variables= get_formal_data(myBiomodModelOut,'expl.var.names')[c(1,2,3,4,5,6,7)], # We specify the variables we want to display
                                  do.bivariate = FALSE,
                                  col = c("red"),
                                  legend = TRUE,
                                  data_species = get_formal_data(myBiomodModelOut,'resp.var'))
```


```{r}
# plot 3D response plots
## Here just for one model

myRespPlot3D <- response.plot2(models  = myModel[1],
                               Data = get_formal_data(myBiomodModelOut,'expl.var'), 
                               show.variables = get_formal_data(myBiomodModelOut,'expl.var.names')[c(1,2,3,4,5,6)],
                               do.bivariate = TRUE,
                               fixed.var.metric = 'median',
                               data_species = get_formal_data(myBiomodModelOut,'resp.var'),
                               legend = F,
                               display_title = FALSE)
```


```{r}
# Click on Zoom to see the enlarged plot
#dev.off() #Delete the plots

```





### Obtaining Predictive Habitat Maps

```{r}
library(raster)
library(rasterVis)
```


Plot of probability of presence of the species

```{r}
# Load the corresponding raster
La_proj_current <- raster("./La/proj_current/proj_current_La.grd")

# Plot
plot(La_proj_current)

# Set color palette
blues <- brewer.pal('Blues', n = 5)
reds <- rev(brewer.pal('Reds', n = 5))

# Combinacion de colores
myTheme <- rasterTheme(region = c(blues, reds))

# Plot
levelplot(La_proj_current, par.settings = BuRdTheme, margin = FALSE, main = expression("Probability of presence | La"))
```



### Predicted habitat plot 

PRESENCE 1 ABSENCE 0 NECESSARY CUT-OFF THRESHOLD POINT DEPENDS ON TECHNIQUE SO AS NOT TO LOSE PREDICTIVE CAPACITY

```{r}
# Load the corresponding raster
La_binary <- raster("./La/proj_current/proj_current_La_ensemble_TSSbin.grd")

#plot
plot(La_binary)

# Alphabetical order
La.bn <- as.factor(La_binary)

# Add a column with the new attribute
rat <- levels(La.bn)[[1]]
rat[["La.bn"]] <- c("Unsuitable", "Suitable")
levels(La.bn) <- rat
```


```{r}
#plot
La.hab <- levelplot(La.bn, col.regions=rev(terrain.colors(2)), xlab="", ylab="", main = expression("Habitat Suitability Index | La"))

plot(La.hab)
```

```{r}

# save as geotiff
if (require(rgdal)) {
   La <- writeRaster(La_binary, filename="./La/proj_current/proj_current_La_ensemble_TSSbin.tif", format="GTiff", overwrite=TRUE)
}
```


