---
title: "Conflict Maps (I)"
author: "Jessica Bernal"
date: "2022-08-14"
output: html_document
---

### 6. Mapping Conflict zones. Preparing variables.

```{r}
#setwd if neccesary. Here we only check it, as it has been previously defined for the project
getwd()
```

Loading the libraries to use,

```{r}
library(raster)
library(rgdal)
library(spatstat)
library(spatialEco)
library(usdm) # Uncertainty analysis for species distribution models
library(ggplot2) # Powerful graphics language for creating elegant and complex plots
```


Loading the data and creating the paths to use.


```{r}
### Conflict species rasters path
path2layers<-"./Proyecto/Conflict_species"

### Results path 
path2work1<-"./Proyecto/Results/"

### Loading species rasters
CONF_SP <- stack(list.files(path=path2layers,pattern='',full.names=TRUE))
names(CONF_SP)
```

Load the variables of Population density, Land cover and Livestock separatedly

```{r}
### Population density
VAR_pop <- raster("./Population_density/GHSL_JRC_P2020/GHS_POP_E2020_GLOBE_R2022A_54009_100_V1_0_R12_C22.tif")

plot(VAR_pop)
```
```{r}
### Land cover map
VAR_land <- raster("./Land_cover/E020S20_PROBAV_LC100_global_v3.0.1_2019-nrt_Discrete-Classification-map_EPSG-4326.tif")

plot(VAR_land)
```
```{r}
### Livestock - buffalo
VAR_ls_buffalo <- raster("./Livestock/buffaloes/5_Bf_2010_Da.tif")

plot(VAR_ls_buffalo)
```
```{r}
### Livestock - cattle
VAR_ls_cattle <- raster("./Livestock/cattle/5_Ct_2010_Da.tif")

plot(VAR_ls_cattle)
```
```{r}
### Livestock - chickens
VAR_ls_chicken <- raster("./Livestock/chickens/5_Ch_2010_Da.tif")

plot(VAR_ls_chicken)
```
```{r}
### Livestock - ducks
VAR_ls_duck <- raster("./Livestock/ducks/5_Dk_2010_Da.tif")

plot(VAR_ls_duck)
```

```{r}
### Livestock - goats
VAR_ls_goat <- raster("./Livestock/goats/5_Gt_2010_Da.tif")

plot(VAR_ls_goat)
```
```{r}
### Livestock - horses
VAR_ls_horse <- raster("./Livestock/horses/5_Ho_2010_Da.tif")

plot(VAR_ls_horse)
```
```{r}
### Livestock - pigs
VAR_ls_pig <- raster("./Livestock/pigs/5_Pg_2010_Da.tif")

plot(VAR_ls_pig)
```
```{r}
### Livestock - sheep
VAR_ls_sheep <- raster("./Livestock/sheep/5_Sh_2010_Da.tif")

plot(VAR_ls_sheep)
```

```{r}
extent(CONF_SP)
extent(VAR_pop)
extent(VAR_land)
extent(VAR_ls_buffalo)
extent(VAR_ls_cattle)
extent(VAR_ls_chicken)
extent(VAR_ls_duck)
extent(VAR_ls_goat)
extent(VAR_ls_horse)
extent(VAR_ls_pig)
extent(VAR_ls_sheep)
```

```{r}
xres(CONF_SP)
xres(VAR_pop)
xres(VAR_land)
xres(VAR_ls_buffalo)
xres(VAR_ls_cattle)
xres(VAR_ls_chicken)
xres(VAR_ls_duck)
xres(VAR_ls_goat)
xres(VAR_ls_horse)
xres(VAR_ls_pig)
xres(VAR_ls_sheep)
```
```{r}
crs(CONF_SP)
crs(VAR_pop)
crs(VAR_land)
crs(VAR_ls_buffalo)
crs(VAR_ls_cattle)
crs(VAR_ls_chicken)
crs(VAR_ls_duck)
crs(VAR_ls_goat)
crs(VAR_ls_horse)
crs(VAR_ls_pig)
crs(VAR_ls_sheep)
```
```{r}
# Transform CRS if necessary
VAR_pop <- projectRaster(VAR_pop, crs = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")

```
```{r}
memory.limit()
```

```{r}
memory.limit(36000)
```

```{r}
VAR_pop <- projectRaster(VAR_pop, crs = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
```


```{r}

# Assign coordinate system if necessary
#crs(VAR_pop) <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"
"
```

```{r}
crs(VAR_pop) 
```

Crop rasters with AOI

```{r}
### AOI path
PN_shp <- readOGR("./PN_Limpopo/limites_limpopo/PN_Limpopo_30km.shp")

### crs transform
PN_shp <- spTransform(PN_shp, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
```
```{r}
crs(PN_shp)
```

Crop and mask with AOI

```{r}
## Population density

VAR_pop <- crop (VAR_pop, extent(PN_shp))
VAR_pop <- mask(VAR_pop, PN_shp)

## Check that it worked
plot(VAR_pop)
plot(PN_shp, add=TRUE, lwd=2)
```

```{r}
## Land cover

VAR_land <- crop (VAR_land, extent(PN_shp))
VAR_land <- mask(VAR_land, PN_shp)

## Check that it worked
plot(VAR_land)
plot(PN_shp, add=TRUE, lwd=2)
```

```{r}
## Livestock buffaloes

VAR_ls_buffalo <- crop (VAR_ls_buffalo, extent(PN_shp))
VAR_ls_buffalo <- mask(VAR_ls_buffalo, PN_shp)

## Check that it worked
plot(VAR_ls_buffalo)
plot(PN_shp, add=TRUE, lwd=2)
```

```{r}
## Livestock catlle

VAR_ls_cattle <- crop (VAR_ls_cattle, extent(PN_shp))
VAR_ls_cattle <- mask(VAR_ls_cattle, PN_shp)

## Check that it worked
plot(VAR_ls_cattle)
plot(PN_shp, add=TRUE, lwd=2)
```
```{r}
## Livestock chicken

VAR_ls_chicken <- crop (VAR_ls_chicken, extent(PN_shp))
VAR_ls_chicken <- mask(VAR_ls_chicken, PN_shp)

## Check that it worked
plot(VAR_ls_chicken)
plot(PN_shp, add=TRUE, lwd=2)
```
```{r}
## Livestock duck

VAR_ls_duck <- crop (VAR_ls_duck, extent(PN_shp))
VAR_ls_duck <- mask(VAR_ls_duck, PN_shp)

## Check that it worked
plot(VAR_ls_duck)
plot(PN_shp, add=TRUE, lwd=2)
```
```{r}
## Livestock goat

VAR_ls_goat <- crop (VAR_ls_goat, extent(PN_shp))
VAR_ls_goat <- mask(VAR_ls_goat, PN_shp)

## Check that it worked
plot(VAR_ls_goat)
plot(PN_shp, add=TRUE, lwd=2)
```

```{r}
## Livestock horse

VAR_ls_horse <- crop (VAR_ls_horse, extent(PN_shp))
VAR_ls_horse <- mask(VAR_ls_horse, PN_shp)

## Check that it worked
plot(VAR_ls_horse)
plot(PN_shp, add=TRUE, lwd=2)
```

```{r}
## Livestock pig

VAR_ls_pig <- crop (VAR_ls_pig, extent(PN_shp))
VAR_ls_pig <- mask(VAR_ls_pig, PN_shp)

## Check that it worked
plot(VAR_ls_pig)
plot(PN_shp, add=TRUE, lwd=2)
```

```{r}
## Livestock sheep

VAR_ls_sheep <- crop (VAR_ls_sheep, extent(PN_shp))
VAR_ls_sheep <- mask(VAR_ls_sheep, PN_shp)

## Check that it worked
plot(VAR_ls_sheep)
plot(PN_shp, add=TRUE, lwd=2)
```





We match extension and resolution of the variables

```{r}
dim(CONF_SP)
dim(VAR_pop)
dim(VAR_land)
dim(VAR_ls_buffalo)
dim(VAR_ls_cattle)
dim(VAR_ls_chicken)
dim(VAR_ls_duck)
dim(VAR_ls_goat)
dim(VAR_ls_horse)
dim(VAR_ls_pig)
dim(VAR_ls_sheep)
```

```{r}
### Match rasters
VAR_pop_new<-resample(VAR_pop,CONF_SP)
dim(VAR_pop_new)<-c(275,230)
plot(VAR_pop_new)
```


```{r}
VAR_land_new<-resample(VAR_land,CONF_SP)
dim(VAR_land_new)<-c(275,230)
plot(VAR_land_new)
```



```{r}
VAR_ls_bf_new<-resample(VAR_ls_buffalo,CONF_SP)
dim(VAR_ls_bf_new)<-c(275,230)
plot(VAR_ls_bf_new)
```

```{r}
VAR_ls_ct_new<-resample(VAR_ls_cattle,CONF_SP)
dim(VAR_ls_ct_new)<-c(275,230)
plot(VAR_ls_ct_new)
```

```{r}
VAR_ls_ch_new<-resample(VAR_ls_chicken,CONF_SP)
dim(VAR_ls_ch_new)<-c(275,230)
plot(VAR_ls_ch_new)
```
```{r}
VAR_ls_dk_new<-resample(VAR_ls_duck,CONF_SP)
dim(VAR_ls_dk_new)<-c(275,230)
plot(VAR_ls_dk_new)
```
```{r}
VAR_ls_gt_new<-resample(VAR_ls_goat,CONF_SP)
dim(VAR_ls_gt_new)<-c(275,230)
plot(VAR_ls_gt_new)
```
```{r}
VAR_ls_ho_new<-resample(VAR_ls_horse,CONF_SP)
dim(VAR_ls_ho_new)<-c(275,230)
plot(VAR_ls_ho_new)
```
```{r}
VAR_ls_pg_new<-resample(VAR_ls_pig,CONF_SP)
dim(VAR_ls_pg_new)<-c(275,230)
plot(VAR_ls_pg_new)
```

```{r}
VAR_ls_sh_new<-resample(VAR_ls_sheep,CONF_SP)
dim(VAR_ls_sh_new)<-c(275,230)
plot(VAR_ls_sh_new)
```




```{r}

extent(CONF_SP)
extent(VAR_pop_new)
extent(VAR_land_new)
extent(VAR_ls_bf_new)
extent(VAR_ls_ct_new)
extent(VAR_ls_ch_new)
extent(VAR_ls_dk_new)
extent(VAR_ls_gt_new)
extent(VAR_ls_ho_new)
extent(VAR_ls_pg_new)
extent(VAR_ls_sh_new)
```

```{r}
xres(CONF_SP)
xres(VAR_pop_new)
xres(VAR_land_new)
xres(VAR_ls_bf_new)
xres(VAR_ls_ct_new)
xres(VAR_ls_ch_new)
xres(VAR_ls_dk_new)
xres(VAR_ls_gt_new)
xres(VAR_ls_pg_new)
xres(VAR_ls_sh_new)
```


```{r}
crs(CONF_SP)
crs(VAR_pop_new)
crs(VAR_land_new)
crs(VAR_ls_bf_new)
crs(VAR_ls_ct_new)
crs(VAR_ls_ch_new)
crs(VAR_ls_dk_new)
crs(VAR_ls_gt_new)
crs(VAR_ls_pg_new)
crs(VAR_ls_sh_new)
```


```{r}
(CONF_SP)
(VAR_pop_new)
(VAR_land_new)
(VAR_ls_bf_new)
(VAR_ls_ct_new)
(VAR_ls_ch_new)
(VAR_ls_dk_new)
(VAR_ls_gt_new)
(VAR_ls_pg_new)
(VAR_ls_sh_new)
```



```{r}
### Load variables
VAR_PN <- stack(c(CONF_SP, VAR_pop_new, VAR_land_new, VAR_ls_bf_new, VAR_ls_ct_new, VAR_ls_ch_new, VAR_ls_dk_new, VAR_ls_gt_new, VAR_ls_pg_new, VAR_ls_sh_new))
plot(VAR_PN)
```
```{r}
names(VAR_PN)
```
```{r}
names(VAR_PN) <- c('Ha_PresentEnsemble', 'La_PresentEnsemble', 'Sc_PresentEnsemble', 'Pop_Density_2020', 'Land_Cover_2019', 'Ls_Bf_2010', 'Ls_Ct_2010', 'Ls_Ch_2010', 'Ls_Dk_2010', 'Ls_Gt_2010', 'Ls_Pg_2010', 'Ls_Sh_2010')

names(VAR_PN)
```

```{r}
### Save variables
writeRaster(x=VAR_PN, path2work1, names(VAR_PN), bylayer=TRUE, format="ascii", overwrite=TRUE)
```

We check the results in QGIS and proceed to normalization an reclassification of the variables

